Econ Analysis

### 2020-04-28 Econ/Econ_Analysis.Rmd
Updated Econ after reviewers

EEZIDs_List <- read_excel("./Econ/Data/Spatial/Updated_EEZList_17June2016.xlsx")
EEZ_CellID <- read_excel("./Econ/Data/Spatial/EEZ_CellID.xlsx")
colnames(EEZ_CellID) <- c("EEZID","INDEX")

# DBEM Coordinate system
Coor <- fread("./Econ/Data/Spatial/Lon_Lat_DBEM.txt",header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

@@

# From (/Users/jpalacios/Documents/UBC/Oceans_Project/Distribution/By Cell/By_Cell_Analysis.Rmd)
Raw_DBEM_Results <- data.table::fread("./Ecol/Data/Distribution/Raw_Data.csv") %>% 
  filter(Species != "Anoplopoma fimbria") #Remove

@@

####_____ IPHC Data ####
# Data from IPHC https://iphc.int/data/commercialresearch-landing-weights-datasets
# Data in metric tonne
IPHC_Landings <- read_excel("./Econ/Data/IPHC/iphc-19912017-commres.xlsx", 
                                    col_types = c("numeric", "text", "text", 
                                                  "numeric", "numeric", "numeric", 
                                                  "text"), skip = 2)
# Clean  landings

@@

# 
GoM_Landings <- read.csv("./Econ/Data/SAU_Landings/GoM/SAU MEOW 55 v47-1.csv")

@@

Fishing_Cost <- read_excel("./Econ/Data/Economic/FishingCost_USA_Canada_20Feb2017.xlsx", 
                           col_types = c("text", "text", "text", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric"), 
                           skip = 1)

@@

Ex_Vessel <- read.delim("./Econ/Data/Economic/Price_US_Canada_5spp.csv")

@@

>>>>

EEZIDs_List <- read_excel("~/Documents/UBC/OC_Special_Issue/Econ/Data/Spatial/Updated_EEZList_17June2016.xlsx")
EEZ_CellID <- read_excel("~/Documents/UBC/OC_Special_Issue/Econ/Data/Spatial/EEZ_CellID.xlsx")
colnames(EEZ_CellID) <- c("EEZID","INDEX")

# DBEM Coordinate system
Coor <- fread("~/Documents/UBC/OC_Special_Issue/Econ/Data/Spatial/Lon_Lat_DBEM.txt",header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

@@

# From (/Users/jpalacios/Documents/UBC/Oceans_Project/Distribution/By Cell/By_Cell_Analysis.Rmd)
Raw_DBEM_Results <- data.table::fread("~/Documents/UBC/OC_Special_Issue/Ecol/Data/Distribution/Raw_Data.csv") %>% 
  filter(Species != "Anoplopoma fimbria") #Remove

@@

####_____ IPHC Data ####
# Data from IPHC https://iphc.int/data/commercialresearch-landing-weights-datasets
# Data in metric tonne
IPHC_Landings <- read_excel("./Data_Econ/IPHC/iphc-19912017-commres.xlsx", 
                                    col_types = c("numeric", "text", "text", 
                                                  "numeric", "numeric", "numeric", 
                                                  "text"), skip = 2)
# Clean  landings

@@

# 
GoM_Landings <- read.csv("~/Documents/UBC/OC_Special_Issue/Econ/Data/SAU_Landings/GoM/SAU MEOW 55 v47-1.csv")

@@

Fishing_Cost <- read_excel("~/Documents/UBC/OC_Special_Issue/Econ/Data/Economic/FishingCost_USA_Canada_20Feb2017.xlsx", 
                           col_types = c("text", "text", "text", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric", "numeric", 
                                         "numeric", "numeric"), 
                           skip = 1)

@@

Ex_Vessel <- read.delim("~/Documents/UBC/OC_Special_Issue/Econ/Data/Economic/Price_US_Canada_5spp.csv")

@@

# Sensitivity Analysis

## Multiple fishing cost and price

Run the analysis with increses in prices and costs of 5%, 10% and 15%

```{r fishing_cost_price, eval = F, echo = F}

## -------------------- ##
#  Create sensitivity data
## -------------------- ##

# random_price <- runif(3,1.05,1.5)
# [1] A) 1.362173 B) 1.425854 C) 1.061325

# Ex vessel sensitivity prices
Sen_ex_vessel_data <- Clean_Ex_Vessel %>% 
  mutate(
    Level_A = Price_perTon*random_price[1],
    Level_B = Price_perTon*random_price[2],
    Level_C = Price_perTon*random_price[3],
    Original = Price_perTon
  ) %>% 
  gather("Level","Sen_Price",Level_A:Original)

# random_cost <- runif(3,1.05,1.5)
# [1] A) 1.311245 B) 1.427684 c) 1.323787

# Profit sensitivity
Sen_cost_data <- Clean_Fishing_Cost %>%
  mutate(
    Level_A = Cost_perTon_2010*random_cost[1],
    Level_B = Cost_perTon_2010*random_cost[2],
    Level_C = Cost_perTon_2010*random_cost[3],
    Original = Cost_perTon_2010
  ) %>% 
  gather("Level","Sen_Cost",Level_A:Original)

## -------------------- ##
# Estimate Profits 
## -------------------- ##

Sen_Profit_Data_Weight <- Landings_Data %>% 
  left_join(Sen_cost_data,
            by = c("gear_type","EEZ")) %>% # Joins Fishing cost Data
  left_join(Sen_ex_vessel_data,
            by = c("Species","EEZ","Level")) %>% 
  mutate(
    Sen_Profit = Sen_Price-Sen_Cost
  )

Sen_Profit_Data_Paper <- Sen_Profit_Data_Weight %>% 
  mutate_at(vars(Sen_Cost,Sen_Price,Sen_Profit), # note these are totals and not per tone
            funs(round((.*Avrg_Landings)/1000000,2)) # million USD
            ) %>% 
  group_by(EEZ,Species,Level) %>% 
  summarise_if(is.numeric,sum,na.rm=T) %>% 
  filter(Species != "Melanogrammus aeglefinus")

# Project profits

Sen_Projected_Profits <- Sen_Profit_Data_Weight %>%
  left_join(MCP_Change,
            by = c("Species","EEZ")
            ) %>%
  arrange(Year) %>% # so the bumber of years match the actual year
  # bind_cols(Years) %>% 
  mutate(
    Future_Landings = Avrg_Landings*Prop_Change,
    Profit = Future_Landings*Sen_Profit,
    # Discounted_Profit = Profit/(1*(1 +dr)^Years_Dis)
  ) %>%
  group_by(Species,EEZ,Model,RCP,Year,Period,Level) %>% # Summ all profits per gear
  summarise(
    Non_Discounted_Profits = sum(Profit)#,
    # Discounted_Profits = sum(Discounted_Profit)
  ) %>%
  filter(Period != "Other_Years") %>%
  ungroup() %>%
  # group by time period
  group_by(Species,Model,Period,EEZ,RCP,Level) %>%
  summarise_if(is.numeric,mean,na.rm=T) %>% 
  # group by ESM
  group_by(Species,Period,EEZ,RCP,Level) %>%
  summarise_if(is.numeric,mean,na.rm=T) %>% 
  filter(RCP == "High Emission",
         Period != "Begining",
         Species != "Melanogrammus aeglefinus") %>% 
  mutate(
    Plot_Labels = gsub("_"," ",Level) # just to fix plot
  )

# Bar graph for sensitivity

ggplot(Sen_Projected_Profits) +
  geom_bar(
    aes(
      x = reorder(Period,desc(Period)),
      y = Non_Discounted_Profits/1000000,
      fill = Plot_Labels
    ),
    position="dodge",
    stat = "identity"
  ) +
  labs(x = "",
       y = "Profits (million USD)"
       ) +
  # scale_fill_discrete("Scenario",values=)+
  scale_fill_brewer("Scenario", type='div', palette="Set3") +
  facet_wrap(~EEZ+Species,
             scales = "free")

ggsave("sen_econ.png",
         plot = last_plot(),
         width = 10,
         height = 8,
         units = "in"#,
         # path = Fig_path_out
         )
